\input{title.tex}

### 2022-04-11 docs/papers/01_manuscript/01_introduction.tex
docs(introduction-refactor): introduction re-write according to Shuai Wang's suggestions. 03_causes changed by grammly, drafts added. metadata title and version updated
Now, we focus more on water governance in the yellow river basin, to make reader curious about how institution effect on that.

BREAKING CHANGE: Writing logic totally changed.

% 水竞争的重要性
Widespread freshwater scarcity and overuse, resulting in systematic risks to economies, societies, and ecosystems globally, are critical environmental challenges to sustainable development \cite{distefano2017, dolan2021, xu2020b, mekonnen2016}.
With steadily increasing demand, competition for water is an urgent problem in water governance, where policies often lead to long-term changes in human–water relationships and the redistribution of benefits \cite{gleick2010, ziolkowska2016, wang2019d}.
Despite governments worldwide trying to resolve competition for water through deliberate institutions at large river basins, cascading effects of these initiatives are poorly understood
\cite{giuliani2013, falkenmark2019, jaeger2019}.

% 制度的意义
Institutions (such as policies, laws, and norms) can influence regional sustainability by changing the structure of the coupled system, including interplays between social actors, ecological units, or between social and ecological system elements \cite{young2008,cumming2020b,lien2020, bodin2017b}.
Understanding the complex interlinkages is crucial for developing strategies to effectively manage natural resources and enhance the resilience of social-ecological systems (SES) \cite{kluger2020}.
Effective (“matched”) institutions operate at appropriate spatial, temporal, and functional scales to manage and balance these different relationships and interactions, therefore, supporting (but do not guarantee) sustainability of SES \cite{epstein2015, wang2019d}.
Water governance tends to shift to institutional solutions within a complex basinal system, where societal drivers impact through water use and related technical interventions \cite{fischer2020}.
Some institutional shifts have desirable water governance outcomes (e.g., the Ecological Water Diversion Project in Heihe River Basin, China \cite{wang2019d} and collaborative water governance systems in Europe \cite{green2013}).
However, shifting institutions in a large, complex river basin may create or destroy hundreds of different connections, where matched human-water relationships are not ubiquitous.
%TODO 这里应该梳理 Institutional shift --> structures / building blocks.
Therefore, despite widespread recognition of the rising importance of institutions as an approach to water sustainable use within large river (especially transboundary river) basins, broader cascading effects of these changes are still in open discussion \cite{agrawal2003, persha2011, agrawal2001}.
% TODO 根据武老师的意见，这个KG应该再明确

%! 这一段梳理一下 building blocks 的作用机制
% Among

% 黄河的制度介绍
Supporting $35.63\%$ irrigation and $xx\%$ population with only $2.66\%$ of water resources in China (data from \href{http://www.yrcc.gov.cn}{http://www.yrcc.gov.cn}, last access: 28 February 2021), the overburdened Yellow River (YR) once dried up in consecutive years.
For that, as a pioneer in water governance shifting in China, the YRB started to explore the initial water allocating scheme in the 1970s, then found a successful solution of dring-up in 1998, and promoted entirely since 2008.
The institutional shifts in 1987 (known as the ``87 Water Allocation Scheme, 87-WAS'') and 1998 (Unified Basinal Regulation, 98-UBR) were two widely recognized milestones of water governance.
Before the remarkable achievement in the restoration of river depletion since the 98-UBR, on the contrary, the first temptation to restrict water uses in the 87-WAS was recognized as a not fulfilling expectation of institutional shift.
Until the 87-WAS, stakeholders have free access to the YR water resources, with geographic and temporal differences between freshwater demand and availability.
As the mismatch between demands and supply kept increasing, national authorities proposed in 87-WAS allocating specific water quotas between $10$ provinces (or regions) along the YR basin.
However, this controversial scheme helped little in turning water depletion around until a different strategy expanded the responsibilities of basinal authorities in integrated water management in 1998 and started progress in the restoration.

% 我们的研究
The shifts in the water governance institution of YR refactored the interplays between humans and water with long-term cascading effects, leaving two quasi-natural experiments for understanding the interactions.
By Differenced Synthetic Control method, our analysis of the net effects of institutional shift shows that institutional mismatches contributed to the acceleration of water withdrawals for resource users after 98-WAS.
As few large river basins have experienced such radical structural changes several times, our quantitative analysis of institutional shifts in the YR induces a valuable understanding of water sustainable governance by decoupling natural and human interferences.
%! 这还要继续凝练我们的贡献
By highlighting potential concerns for ecosystem collapse under structural mismatches, our findings align with the urgent calls for a more dynamic design for water governance institutions to achieve sustainability.

>>>>

% 水竞争的重要性
Widespread freshwater scarcity and overuse, resulting in systematic risks to economies, societies, and ecosystems globally, are critical environmental challenges to sustainability \cite{distefano2017, dolan2021, xu2020b, mekonnen2016}.
With steadily increasing demand, competition for water causes depletion of freshwater globally and calls for an urgent transformation of governance system by considering water use conservation \cite{gleick2010, ziolkowska2016, wang2019d}.
Despite worldwide trying to govern water, however, degredation of large river basins is not easily reversable because of few alignments between practice and theory in successful water governance cases.
\cite{giuliani2013, falkenmark2019, jaeger2019}.

% 黄河的介绍
The Yellow River Basin (YRB), the fifth large river worldwide, is known for irreplacable role in the social-economic development of China, and thus also drastically interferend by anthropotic stress.
Supporting $35.63\%$ irrigation and $30\%$ population with only $2.66\%$ water resources of China (data from \href{http://www.yrcc.gov.cn}{http://www.yrcc.gov.cn}, last access: 28 February 2021), the overburdened Yellow River dried up in consecutive years, resulting huge ecological, economic and social crisis (e.g., wetland shrink, agriculture reduction, and scramble for water).
Intense water use, accounting for about $80\%$ of Yellow River surface runoff in 1980s, was remarked as the major reason of the degradation.
Furthermore, human interferences such as soil conservation and water conservancy project boosted water withdraw then stressed the water scarcity of the Yellow River.
To balance between ecological and developing demands in such a human-dominated basin, therefore, is not only a problem for China in water governance throughout, but also for large rivers worldwide.

% 第三段 过去管理的经验，断流的严重程度，连续多少没有断流，少有的解决断流的大河。治理成功的流域？分析背后的机制，基础。往用水上靠。
In China, several ambitious water management practices were proposed since the last centery, such as reservoir regulating, South-to-north Water Diversion Project (WDP), Water Allocation Scheme since 1987 (87-WAS), and the Unified Basinal Regulation since 1998 (98-UBR), to solve the water stress of the YRB.
Contributed by those efforts, ecological restoration of wetland and estuary delta in the YRB without drying-up for over $20$ years is widely considered as a achievement in large river management.
Different from engineering of WDP provides further water supply or reservoir matches water supply and demands, however, institutional strategies like the 87-WAS (assign water quotas for provinces in the YRB) and the 98-UBR (the provinces had to be allowed for using water by the Yellow River Conservancy Commission, YRCC) mainly focused on limiting demands of water use.
Such institutions (policies, laws, and norms) can influence regional sustainability by changing the structure of the coupled human and natural system, including interplays between social actors, ecological units, or between social and ecological system elements \cite{young2008,cumming2020b,lien2020, bodin2017b}.
Understanding the complex interlinkages is crucial for developing strategies to effectively manage natural resources and enhance the resilience of social-ecological systems (SES) \cite{kluger2020}.
However, while effects of engineering solutions are well evaluated and quantified beforehand, there are few attempts on assessment of institutional contributions to an successful water governance.

% 第四段 目标 制度变化。是不是制度变化产生的，如果是的话怎么影响。
Despite widespread recognition of the rising importance of institutions as an approach to water sustainable use within large river (especially transboundary river) basins, their specific effects are still in open discussion \cite{agrawal2003, persha2011, agrawal2001}.
Effective (``matched'' or ``fit'') institutions operate at appropriate spatial, temporal, and functional scales to manage and balance different relationships and interactions between human and water system, therefore, supporting (but do not guarantee) sustainability of SES \cite{epstein2015, wang2019d}.
Some institutional shifts have desirable water governance outcomes (e.g., the Ecological Water Diversion Project in Heihe River Basin, China \cite{wang2019d} and collaborative water governance systems in Europe \cite{green2013}).
However, shifting institutions in a large, complex river basin may create or destroy hundreds of different connections, where matched human-water relationships are not ubiquitous.
How did and how important the institutional shifts played an role in the water governance achievement of the YRB, therefore, are still uncertain.
Here, we quantitatively analyzed impacts of two significant institutional shifts (the 87-WAS and the 98-UBR) on water uses of the YRB.
By linking structures and outcomes within the coupled human and natural system, we further explored mechanisms of institutional effects in the YRB, for deeper understanding of institutions role in water governance worldwide.

### 2022-04-11 docs/papers/01_manuscript/03_causes.tex
docs(introduction-refactor): introduction re-write according to Shuai Wang's suggestions. 03_causes changed by grammly, drafts added. metadata title and version updated
Now, we focus more on water governance in the yellow river basin, to make reader curious about how institution effect on that.

BREAKING CHANGE: Writing logic totally changed.

% \MakeUppercase{\subsection{Mechanism of institutional shift effects}}
\subsection{Mechanism of institutional shift effects}
\label{result-2}
% 结果2部分：
% 1. 加速用水增长的结果 for each province.
% 2. Potential benefits for each stakeholder (economic model - assumption & results), and social-ecological fit holistically. 结合经济模型
% 3. Why some building blocks fits for good outcomes. 结合 building block

Differences between stakeholders in responses to institutional shifts are vital to understanding the mechanism between structures and outcomes.
Our results show that the proportion of accelerated water use in each province after the decade of 87-WAS (the proportion of actual water use exceeding the predicted water use by the model) has a significant correlation ($p<0.05$, see \textbf{Methods}) to the Yellow River water use in each province (Figure~\ref{upset}A).
Furthermore, while no evident impacts for most provinces (no more than $10\%$ differences, the apparent acceleration effects were only prominent in the big water-using provinces (e.g., Neimeng, Henan, and Shandong. Figure~\ref{upset}B).
In particular, Neimeng and Shandong, both provinces that exceeded the prescribed water uses of the 87-WAS, used $44.25\%$ and $25.69\%$ more water uses than the prediction from 1987 to 1998, respectively.
Furthermore, the satisfaction of each province with the water allocation stipulated by the 87-WAS (expressed by the difference between the actual water allocation of each province and the expected planning value) is not significantly correlated with the acceleration.
By contrast, after the 98-UBR, with the exception of Shaanxi (which has always been abundant in water quota) had evident ($17.53\%$) increasd water use, almost all provinces have seen significant declines in water use ($-12.5\%$ on average).
Neither the satisfaction nor Yellow River water use, however, have correlation with the declines.

\begin{figure*}[!h]
    \centering
    \includegraphics[width=32pc]{outputs/upset.jpg}
    \caption{
        \textbf{A.} The partial correlation coefficient between wate uses (WU) of Yellow River (YR), unsatisfied ratio (compared with requirements in water plan and supply in the 87-WAS), and the average accelerated ratio.
        \textbf{B.} Average accelerated ratio of water uses for each province in the YRB during the decade after 87-WAS (from 1987 to 1998).
        \textbf{Mian users:} Major water consumption provinces (over the median).
        \textbf{Overused:} violate the 87-WAS in average water uses.
    }
    \label{upset}
\end{figure*}

>>>>

% \MakeUppercase{\subsection{Mechanism of institutional shift effects}}
\subsection{Mechanism of institutional effects}
\label{result-2}
% 结果2部分：
% 1. 加速用水增长的结果 for each province.
% 2. Potential benefits for each stakeholder (economic model - assumption & results), and social-ecological fit holistically. 结合经济模型
% 3. Why some building blocks fits for good outcomes. 结合 building block

Differences between stakeholders in responses to institutional shifts are vital to understanding the mechanism between structures and outcomes.
Our results show that the proportion of accelerated water use in each province after the decade of 87-WAS (the proportion of actual water use exceeding the predicted water use by the model) has a significant correlation ($p<0.05$, see \textbf{Methods}) to the Yellow River water use in each province (Figure~\ref{upset}A).
Furthermore, while no evident impacts for most provinces (no more than $10\%$ differences, the apparent acceleration effects were only prominent in the big water-using provinces (e.g., Neimeng, Henan, and Shandong. Figure~\ref{upset}B).
In particular, Neimeng and Shandong, both provinces that exceeded the prescribed water uses of the 87-WAS, used $44.25\%$ and $25.69\%$ more water uses than the prediction from 1987 to 1998, respectively.
Furthermore, the satisfaction of each province with the water allocation stipulated by the 87-WAS (expressed by the difference between the actual water allocation of each province and the expected planning value) has little significant correlation to the acceleration.
By contrast, after the 98-UBR, except Shaanxi (which has always been abundant in water quota) had an evident ($17.53\%$) increase in water use, almost all provinces have seen significant declines in water use ($-12.5\%$ on average).
However, neither the satisfaction nor Yellow River water use correlates with the declines.

\begin{figure*}[!h]
    \centering
    \includegraphics[width=32pc]{outputs/upset_87.pdf}
    \caption{
        \textbf{A.} The partial correlation coefficient between wate uses (WU) of Yellow River (YR), unsatisfied ratio (compared with requirements in water plan and supply in the 87-WAS), and the average accelerated ratio.
        \textbf{B.} Average accelerated ratio of water uses for each province in the YRB during the decade after 87-WAS (from 1987 to 1998).
        \textbf{Mian users:} Major water consumption provinces (over the median).
        \textbf{Overused:} violate the 87-WAS in average water uses.
    }
    \label{upset}
\end{figure*}

### 2022-04-11 docs/papers/01_manuscript/drafs.tex
docs(introduction-refactor): introduction re-write according to Shuai Wang's suggestions. 03_causes changed by grammly, drafts added. metadata title and version updated
Now, we focus more on water governance in the yellow river basin, to make reader curious about how institution effect on that.

BREAKING CHANGE: Writing logic totally changed.

>>>>

% 为此，长江流域作为中国水治理转型的先驱者，从上世纪70年代开始探索初步的配水方案，到1998年找到了解决水资源枯竭问题的成功方案，并从2008年开始全面推广。
% 1987年(被称为“87年水分配计划，87- was”)和1998年(统一流域管理，98-UBR)的制度转变是水治理的两个被广泛认可的里程碑。
% 相反，在自98-UBR以来在恢复河流枯竭方面取得显著成就之前，第一个限制87-WAS用水的诱惑被认为是对体制转变的不满足的期望。
% 在87-WAS之前，利益攸关方可以免费使用YR水资源，但淡水需求和可用性之间存在地理和时间差异。
% 由于需求和供应之间的不匹配不断增加，国家当局在87-WAS中建议在YR盆地沿线10个省(或区域)之间分配具体的水配额。
% 然而，这个有争议的计划在扭转水资源枯竭方面帮助不大，直到1998年一项不同的战略扩大了流域当局在综合水资源管理方面的责任，并开始在恢复方面取得进展。
For that, as a pioneer in water governance shifting in China, the YRB started to explore the initial water allocating scheme in the 1970s, then found a successful solution of dring-up in 1998, and promoted entirely since 2008.
The institutional shifts in 1987 (known as the ``87 Water Allocation Scheme, 87-WAS'') and 1998 (Unified Basinal Regulation, 98-UBR) were two widely recognized milestones of water governance.
Before the remarkable achievement in the restoration of river depletion since the 98-UBR, on the contrary, the first temptation to restrict water uses in the 87-WAS was recognized as a not fulfilling expectation of institutional shift.
Until the 87-WAS, stakeholders have free access to the YR water resources, with geographic and temporal differences between freshwater demand and availability.
As the mismatch between demands and supply kept increasing, national authorities proposed in 87-WAS allocating specific water quotas between $10$ provinces (or regions) along the YR basin.
However, this controversial scheme helped little in turning water depletion around until a different strategy expanded the responsibilities of basinal authorities in integrated water management in 1998 and started progress in the restoration.

Water governance tends to shift to institutional solutions within a complex basinal system, where societal drivers impact through water use and related technical interventions \cite{fischer2020}.

% YR水治理机构的转变重构了人与水之间的相互作用，并产生了长期的级联效应，为理解这种相互作用留下了两个准自然实验。
% 利用差分综合控制方法，对制度变迁的净效应进行了分析，结果表明:制度不匹配促进了98-WAS后资源使用者的取水速度。
% 由于很少有大型河流流域经历过如此剧烈的结构变化，我们对年度制度变迁的定量分析，通过将自然和人类干预脱钩，促使我们对水的可持续治理有了宝贵的理解。
% % !这还要继续凝练我们的贡献
% 通过强调结构不匹配下生态系统崩溃的潜在担忧，我们的研究结果与迫切需要为水治理机构提供更动态的设计以实现可持续性的呼吁相一致。
The shifts in the water governance institution of YR refactored the interplays between humans and water with long-term cascading effects, leaving two quasi-natural experiments for understanding the interactions.
By Differenced Synthetic Control method, our analysis of the net effects of institutional shift shows that institutional mismatches contributed to the acceleration of water withdrawals for resource users after 98-WAS.
As few large river basins have experienced such radical structural changes several times, our quantitative analysis of institutional shifts in the YR induces a valuable understanding of water sustainable governance by decoupling natural and human interferences.
%! 这还要继续凝练我们的贡献
By highlighting potential concerns for ecosystem collapse under structural mismatches, our findings align with the urgent calls for a more dynamic design for water governance institutions to achieve sustainability.
